### 2023-04-25 paper/main.tex
tables and stuff in latex

\documentclass[twocolumn]{article}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{tabularx,ragged2e,booktabs,caption}
\usepackage{multirow}
\usepackage{color}
\usepackage[top=1in,bottom=1in,left=1.2in,right=1.2in]{geometry}
\usepackage{hyperref}
\usepackage[small]{titlesec}
\usepackage{cite}

\renewcommand\tabularxcolumn[1]{C{#1}}

\newcommand{\todo}[1]{\textcolor{cyan}{\textbf{TODO:} #1}}

\title{CS378: Final Project - Data Artifacts \\ \url{https://github.com/sarimaleem/cs378_fp}}
\author{Sarim Aleem \\ ska2222}

\date{}

\begin{document}
\maketitle

\begin{abstract}
    \todo{write an abstract describing the core motivation and findings from your project}
\end{abstract}

\section{Introduction (5pt)}

\todo{This section should include three paragraphs. The first paragraph is to
briefly describe task and data. The second paragraph should describe the
results of your analysis, and the third paragraph describing your fix and main
experimental take aways. }

The model that was used was \cite{clark2020electra}

\section{Task/Dataset/Model Description (15pt)}

\todo{Describe the task/dataset/model you are working on. Clearly define your
task with mathematical notations. Describe your learning algorithm. You must
formally specify the loss function, stopping criteria, training data, etc used
for the model you are analyzing in the next section. Remember, every notation
you use must be defined.}

\section{Performance Analysis (25pt)}

The model was trained for 3 iterations on the SNLI training dataset. Each
iteration has 550152 training examples, the batch size for the training examples
was 32. All training was done on google colab in the cloud, and training took
about 25 minutes.

\subsection*{Overall Accuracy Statistics}

An initial evaluation of the model shows that it has an accuracy of 88\%. We
also did further evaluation to show statistics about which labels it
misclassified most using a confusion matrix.

\begin{table}[]
\caption{Development Set Evaluation Metrics}
\centering
\begin{tabular}{|l|l|}
\hline
loss       & 0.315   \\ \hline
accuracy   & 0.8865  \\ \hline
runtime    & 20.2629 \\ \hline
examples/s & 485.715 \\ \hline
steps/s    & 60.751  \\ \hline
\end{tabular}
\caption*{\textit{Model Metrics for 9842 examples}}

\caption{Confusion Matrix}
\begin{tabular}{|c|cccc|c|}
\hline
\multicolumn{1}{|l|}{}         & \multicolumn{4}{c|}{\textbf{Predicted}}                                                                              & \multicolumn{1}{l|}{} \\ \hline
\multirow{4}{*}{\textbf{True}} & \multicolumn{1}{l|}{}               & \multicolumn{1}{c|}{\textit{0}} & \multicolumn{1}{c|}{\textit{1}} & \textit{2} & \textbf{Total}        \\ \cline{2-6} 
                               & \multicolumn{1}{c|}{\textit{0}}     & \multicolumn{1}{c|}{2996}       & \multicolumn{1}{c|}{252}        & 81         & 3329                  \\ \cline{2-6} 
                               & \multicolumn{1}{c|}{\textit{1}}     & \multicolumn{1}{c|}{213}        & \multicolumn{1}{c|}{2787}       & 235        & 3235                  \\ \cline{2-6} 
                               & \multicolumn{1}{c|}{\textit{2}}     & \multicolumn{1}{c|}{79}         & \multicolumn{1}{c|}{257}        & 2942       & 3278                  \\ \hline
\multicolumn{1}{|l|}{}         & \multicolumn{1}{c|}{\textbf{Total}} & \multicolumn{1}{c|}{3288}       & \multicolumn{1}{c|}{3296}       & 3258       & \textbf{9842}         \\ \hline
\end{tabular}
\caption*{\textit{ 0=Entailment, 1=Neutral, 2=Contradiction }}

\caption{Percent mispredictions}
\begin{tabular}{|l|cccc|}
\hline
                               & \multicolumn{4}{c|}{\textbf{Predicted}}                                                                          \\ \hline
\multirow{4}{*}{\textbf{True}} & \multicolumn{1}{l|}{}           & \multicolumn{1}{c|}{\textit{0}} & \multicolumn{1}{c|}{\textit{1}} & \textit{2} \\ \cline{2-5} 
                               & \multicolumn{1}{c|}{\textit{0}} & \multicolumn{1}{c|}{89.99}      & \multicolumn{1}{c|}{7.78}       & 2.47       \\ \cline{2-5} 
                               & \multicolumn{1}{c|}{\textit{1}} & \multicolumn{1}{c|}{6.39}       & \multicolumn{1}{c|}{86.15}      & 7.16       \\ \cline{2-5} 
                               & \multicolumn{1}{c|}{\textit{2}} & \multicolumn{1}{c|}{2.37}       & \multicolumn{1}{c|}{7.94}       & 89.74      \\ \hline
\end{tabular}
\caption*{\textit{ 0=Entailment, 1=Neutral, 2=Contradiction }}
\end{table}

As can be seen from the table,

\todo{Describe clearly how you conducted your analysis of the model
performance, and report the outcomes. You should provide (1) some examples of
both \textbf{specific errors/behavior} from the model and (2) analysis or
discussion of \textbf{the general class of mistakes the model makes}. For
characterizing the general class, try to come up with rules that can identify a
set of challenging examples (e.g., ``examples containing \emph{not}'') and try
to visualize in charts, graphs, or tables what you believe to be relevant
statistics about the data. For example, how frequent is the class of examples
where such error pattern apply? This part of the report should be at least one
page.}

\section{Describing Your Fix (20pt)}

\todo{You should describe modifications you have made to improve performances.
We will evaluate based on how reasonable is your fix  -- if your fix and
motivation does not link, your modificiation is unreasonable. For example,
perhaps you tried to modify neural network training in a way that is totally
unconnected to your stated goal, or your modification was erroneous. Describe
the baseline model to be compared to your approach as well here. Again, clearly
define all notations. Someone who is reading your report (with reasonable
background, e.g., your classmates in this course) should have a reasonable idea
how to implement it themselves. Describe hyperparameters, including how they
are selected, if you have hyperparameters involved. }

\section{Evaluating Your Fix (25pt)}

\todo{
Your writeup should address how effective is your fix, how broadly applicable
is your fix, etc. Providing a single number (overall accuracy) is necessary but
not sufficient here. For example, if your change made the model better on
challenging NLI examples, you could try to quantify that on one or more slices
of the data, give examples of predictions that are fixed, or even use model
interpretation techniques to try to support claims about how your improved
model is doing its ``reasoning.'' (You can look at the papers listed above to
get a sense of how to do such fine-grained evaluation).  You should report
results from a baseline approach (your initial trained model) as well as your
``best'' method. If doing your own project, baselines such as majority class,
random, or a linear classifier are important to see. \textbf{Ablations}: If you
tried several things, analyze the contribution from each one. These should be
\emph{minimal} changes to the same system; try running things with just one
aspect different in order to assess how important that aspect is. This part of
the report should be at least one page.}

% \todo{Describe the data you use, including how many examples are in the training, development, and test sets. Please also provide shallow statistics of your data, with at least the vocabulary size and instruction length. It is best to report all the statistics, including counts, in a table. Please describe how you compute statistics that can be computed in different ways (i.e., vocabulary size). Describe how you pre-process the data, including how you treat casing, tokenization, unknown words, and anything else that you did to the raw data before providing it to your model. If you are using a subset of the data, explain why and discuss tradeoffs (you will also need to back them with experiments).}

% \section{Implementation Details (3pt)}

% \todo{Briefly describe the implementation details. No need to copy details already specified in the assignment. Include any hyper-parameters the model has. If there are any optimizations that you introduced, this is the place to describe them. Did you have to do any special optimizations to make the method work in reasonable time/memory? Describe it here.}

% \section{}

 \section{Related Work (5pt)}
\todo{Briefly discuss prior research papers related to your approach. This will
likely some papers in the project description document. How is your approach
different from existing studies?}
% \paragraph{Ablations (15pt)}

% \todo{Please report your ablation studies on the development set. Make sure to ablate each feature set independently to show they contribute to your final result.  If a feature set does not improve performance, please do not include it -- it doesn't help. You can still discuss it, but there is really no point in including it in your final model. Reporting the results must be done in a table, and each result must be discussed in the text.}

% \section{Error Analysis (7pt)}

% \todo{Qualitative analysis of selected failure examples. Show and discuss error examples from your development set. You must identify certain classes of errors and use the examples to illustrate them. This is often best to show in a table.}

\section{Conclusion (5pt)}
\todo{Brief conclusion summarizing findings from both numerical results and
qualitative analysis.}

\section*{(Optional) AI Assistance}

\todo{If you have used any AI toolkit (either for writing assistance or code
assistance) for your final project, please describe it here. }

\bibliography{sample}{}
\bibliographystyle{plain}

\end{document}

### 2023-04-25 analyze_output.py

@@ -1,0 +1,54 @@
+import json
+import numpy as np
+from prettytable import PrettyTable
+from tabulate import tabulate
+from sklearn.metrics import confusion_matrix, ConfusionMatrixDisplay
+
+f = open("./eval_output/eval_predictions.jsonl")
+
+
+# {"premise": "Two women are embracing while holding to go packages.", 
+# "hypothesis": "The sisters are hugging goodbye while holding to go packages after just eating lunch.", 
+# "label": 1, 
+# "predicted_scores": [-1.8910499811172485, 3.4238548278808594, -1.99180006980896], 
+# "predicted_label": 1}
+
+mispredict = 0
+i = -1
+
+print("reference")
+print("0:", "Entailment")
+print("1:", "Neutral")
+print("2:", "Contradiction")
+print()
+
+y_true = []
+y_predict = []
+
+# print out wrong lines 
+for (i, line) in enumerate(f):
+    example = json.loads(line)
+    label = example["label"]
+    predict =  example["predicted_label"]
+    y_true.append(label)
+    y_predict.append(predict)
+    if(label != predict):
+        mispredict += 1
+        # print("line:", i)
+        # print("premise:", example["premise"])
+        # print("hypothesis:", example["hypothesis"])
+        # print("label:", example["label"])
+        # print("predicted:", example["predicted_label"])
+        # print()
+
+print(mispredict, "/", i, "mispreditions")
+
+cm = confusion_matrix(y_true=y_true, y_pred=y_predict)
+print(cm)
+
+np.set_printoptions(precision=2)
+percent = cm / cm.astype(np.float32).sum(axis=1) * 100
+table = tabulate(percent)
+print(table)
+# print(np.array_str(percent, precision=2, suppress_small=True))
+

### 2023-04-25 paper/main.tex
some additions to performance analysis

@@
 
-As can be seen from the table,
+In general, the model seems to do well with Separating Entailment and
+Contradiction. For example, it only incorrectly identifies 2.47\% of entailment
+examples as contradiction and 2.37\% of contradiction examples as entailement.
+However, the model struggles more to understand the relationship between,
+entailment and neutral, as well as contradiction and neutral.
 

### 2023-04-25 analyze_output.py

@@ -25,7 +25,7 @@ print()
 y_true = []
 y_predict = []
 
-# print out wrong lines 
+# print out wrong lines for specific analysis
 for (i, line) in enumerate(f):
     example = json.loads(line)
     label = example["label"]
@@ -43,12 +43,13 @@ for (i, line) in enumerate(f):
 
 print(mispredict, "/", i, "mispreditions")
 
+# Confusion Matrix
 cm = confusion_matrix(y_true=y_true, y_pred=y_predict)
 print(cm)
 
+# Confusion Matrix that prints out percentages
 np.set_printoptions(precision=2)
 percent = cm / cm.astype(np.float32).sum(axis=1) * 100
 table = tabulate(percent)
 print(table)
-# print(np.array_str(percent, precision=2, suppress_small=True))
 

### 2023-04-26 paper/main.tex
created program to create contrast data

@@
 
+\subsection{Confusion Matrix Analysis}
 In general, the model seems to do well with Separating Entailment and
@@
 examples as contradiction and 2.37\% of contradiction examples as entailement.
+
 However, the model struggles more to understand the relationship between,
-entailment and neutral, as well as contradiction and neutral.
+entailment and neutral, as well as contradiction and neutral. For example, it
+mispredicted over 7\% of entailments and contradictions as neutrals. As long as 
+mispredicting over 6\% of neutrals as entailments and over 7\% of of neutrals as
+contradictions.
+
+\subsection{Manual Pattern Analysis}
+
+It's difficult to say how the model is evaluating sentences, since it's
+impossible to fully understand its weights. Nevertheless, there do seem to be
+some patterns one evaluating mistakes that the model makes. One pattern that
+seems to happen is that the model seems to be unable to understand different
+words that mean the same thing in a context. For example, the following is an
+error that the model made.
+
+In the example with hypothesis \textit{A man and a woman are looking at produce
+on display} and premise \textit{A man and woman are staring at heads of
+lettuce}. The model is unable to distinguish that lettuce is a type of produce
+and instead makes assumption that they are different, thus leading it to think
+they are contradictions.
+
+\begin{table*}[t]
+\caption{Examples of model misunderstanding words, or not considering their importance}
+\begin{tabular}{|p{4cm}|p{4cm}|p{2.5cm}|p{2.5cm}|}
+\hline
+\textbf{premise}                                                                        & \textbf{Hypothesis}                & \textbf{Gold} & \textbf{Predicted} \\ \hline
+People are throwing tomatoes at each other                                                                    & The people are having a food fight               & entailment                         & contradiction      \\ \hline
+A man and a woman are looking at produce on display.
+& A man and women are staring at heads of lettuce.
+& neutral                            & contradiction      \\ \hline
+Two men sitting on a subway are reading, with coats and scarves on, but have seemed to have lost their pants.
+& The men are wearing pants.
+& contradiction                      & entailment         \\ \hline
+Two men are in an electronics workshop, working on computers or equipment                                     & The men are unaware of what computers are.
+& contradiction                      & neutral            \\ \hline
+\end{tabular}
+\end{table*}
+
+\subsection{Using Contast sets to evaluate the Model}
+subsection{Contrast Sets}
+
+In their article \textit{Evaluating models' local decision boundaries via
+contrast sets} \cite{gardner2020evaluating}, Gardner et al. explain the concept
+of a \textit{contrast set}. A contrast set is a small but meaninful alteration
+in the input data that typically leads to a different gold label.
+
+In order to test if the model is finding artifacts in the data or genuinely
+evaluating it, we annotated 50 different examples in the validation data that
+the model predicted correctly, and then tested to see if the model was able to
+predict them correctly again.
+
+\subsection{Using contast sets to evaluate the model}
 
@@
 discussion of \textbf{the general class of mistakes the model makes}. For
-characterizing the general class, try to come up with rules that can identify a
-set of challenging examples (e.g., ``examples containing \emph{not}'') and try
-to visualize in charts, graphs, or tables what you believe to be relevant
+characterizing the general class, try to come up with rules that can identify
+a set of challenging examples (e.g., ``examples containing \emph{not}'') and
+try to visualize in charts, graphs, or tables what you believe to be relevant
 statistics about the data. For example, how frequent is the class of examples
@@
 page.}
+
+% Describe clearly how you conducted your analysis of the model
+% performance, and report the outcomes. You should provide (1) some examples of
+% both \textbf{specific errors/behavior} from the model and (2) analysis or
+% discussion of \textbf{the general class of mistakes the model makes}. For
+% characterizing the general class, try to come up with rules that can identify a
+% set of challenging examples (e.g., ``examples containing \emph{not}'') and try
+% to visualize in charts, graphs, or tables what you believe to be relevant
+% statistics about the data. For example, how frequent is the class of examples
+% where such error pattern apply? This part of the report should be at least one
+% page.
 
@@
 are selected, if you have hyperparameters involved. }
+
+% You should describe modifications you have made to improve performances.
+% We will evaluate based on how reasonable is your fix  -- if your fix and
+% motivation does not link, your modificiation is unreasonable. For example,
+% perhaps you tried to modify neural network training in a way that is totally
+% unconnected to your stated goal, or your modification was erroneous. Describe
+% the baseline model to be compared to your approach as well here. Again, clearly
+% define all notations. Someone who is reading your report (with reasonable
+% background, e.g., your classmates in this course) should have a reasonable idea
+% how to implement it themselves. Describe hyperparameters, including how they
+% are selected, if you have hyperparameters involved.
 

### 2023-04-26 create_contrast.py

@@ -1,0 +1,32 @@
+import json
+
+in_file = open("./eval_output/eval_predictions.jsonl", "r")
+output = open("./contrast_set.jsonl", "a")
+for (i, line) in enumerate(in_file):
+    if (i > 50):
+        break
+    example = json.loads(line)
+    print("------------------------------------------------------")
+    print("premise:    ", example["premise"])
+    print("hypothesis: ", example["hypothesis"])
+    print("gold label: ", example["label"])
+
+    new_premise  = input("enter new premise (nothing for no change): ").strip()
+    if(new_premise == ""):
+        new_premise = example["premise"]
+    
+    new_hypothesis  = input("enter new hypothesis (nothing for no change): ").strip()
+    if(new_hypothesis == ""):
+        new_hypothesis = example["hypothesis"]
+
+    new_label = input("enter new label (nothing for no change): ").strip()
+    if(new_label == ""):
+        new_label = example["label"]
+    else:
+        new_label = int(new_label)
+
+    new_example = {"premise": new_premise, "hypothesis": new_hypothesis, "label": new_label}
+    json_object = json.dumps(new_example) 
+    output.write(json_object)
+    
+    print("------------------------------------------------------")

### 2023-04-27 paper/main.tex
checkpoint

@@
 In order to test if the model is finding artifacts in the data or genuinely
-evaluating it, we annotated 50 different examples in the validation data that
+evaluating it, we annotated 35 different examples in the validation data that
 the model predicted correctly, and then tested to see if the model was able to
@@
 
-\subsection{Using contast sets to evaluate the model}
+The results of the model showed a significant dip in accuracy, with accuracy
+going down to 77\% from 89\%. The model mispredicted 8/35 examples. One pattern
+that emerged from the mispredictions is that the model fails to see correlations
+between various words and also fails to understand the relationships between
+various words. The following is an example of this phenomenon: \\
 
-\todo{Describe clearly how you conducted your analysis of the model
-performance, and report the outcomes. You should provide (1) some examples of
-both \textbf{specific errors/behavior} from the model and (2) analysis or
-discussion of \textbf{the general class of mistakes the model makes}. For
-characterizing the general class, try to come up with rules that can identify
-a set of challenging examples (e.g., ``examples containing \emph{not}'') and
-try to visualize in charts, graphs, or tables what you believe to be relevant
-statistics about the data. For example, how frequent is the class of examples
-where such error pattern apply? This part of the report should be at least one
-page.}
+\begin{quote}
+premise: Two men on vehicles competing in a race. \\
+hypothesis: Men are riding cars on the street. \\
+label: neutral \\
+predicted: contradiction
+\end{quote}
 
-% Describe clearly how you conducted your analysis of the model
-% performance, and report the outcomes. You should provide (1) some examples of
-% both \textbf{specific errors/behavior} from the model and (2) analysis or
-% discussion of \textbf{the general class of mistakes the model makes}. For
-% characterizing the general class, try to come up with rules that can identify a
-% set of challenging examples (e.g., ``examples containing \emph{not}'') and try
-% to visualize in charts, graphs, or tables what you believe to be relevant
-% statistics about the data. For example, how frequent is the class of examples
-% where such error pattern apply? This part of the report should be at least one
-% page.
+The neural network is essentially unable to distinguish that different words can
+have the same meaning (in this case vehicles and cars).
 
@@
 are selected, if you have hyperparameters involved. }
-
-% You should describe modifications you have made to improve performances.
-% We will evaluate based on how reasonable is your fix  -- if your fix and
-% motivation does not link, your modificiation is unreasonable. For example,
-% perhaps you tried to modify neural network training in a way that is totally
-% unconnected to your stated goal, or your modification was erroneous. Describe
-% the baseline model to be compared to your approach as well here. Again, clearly
-% define all notations. Someone who is reading your report (with reasonable
-% background, e.g., your classmates in this course) should have a reasonable idea
-% how to implement it themselves. Describe hyperparameters, including how they
-% are selected, if you have hyperparameters involved.
 

### 2023-04-27 analyze_contrast.py

@@ -1,0 +1,40 @@
+import json
+import numpy as np
+from prettytable import PrettyTable
+from tabulate import tabulate
+from sklearn.metrics import confusion_matrix, ConfusionMatrixDisplay
+
+f = open("./contrast_output/eval_predictions.jsonl")
+
+
+# {"premise": "Two women are embracing while holding to go packages.", 
+# "hypothesis": "The sisters are hugging goodbye while holding to go packages after just eating lunch.", 
+# "label": 1, 
+# "predicted_scores": [-1.8910499811172485, 3.4238548278808594, -1.99180006980896], 
+# "predicted_label": 1}
+
+mispredict = 0
+i = -1
+
+y_true = []
+y_predict = []
+
+# print out wrong lines for specific analysis
+for (i, line) in enumerate(f):
+    example = json.loads(line)
+    label = example["label"]
+    predict =  example["predicted_label"]
+    y_true.append(label)
+    y_predict.append(predict)
+    if(label != predict):
+        mispredict += 1
+        print("line:", i)
+        print("premise:", example["premise"])
+        print("hypothesis:", example["hypothesis"])
+        print("label:", example["label"])
+        print("predicted:", example["predicted_label"])
+        print()
+        print("----------------------------------------")
+        print()
+
+print(mispredict, "/", i, "mispreditions")

### 2023-04-27 create_contrast.py

@@ -3,6 +3,9 @@ import json
 in_file = open("./eval_output/eval_predictions.jsonl", "r")
 output = open("./contrast_set.jsonl", "a")
 for (i, line) in enumerate(in_file):
+    if (i < 30):
+        continue
+
     if (i > 50):
         break
     example = json.loads(line)
@@ -27,6 +30,6 @@ for (i, line) in enumerate(in_file):
 
     new_example = {"premise": new_premise, "hypothesis": new_hypothesis, "label": new_label}
     json_object = json.dumps(new_example) 
-    output.write(json_object)
+    output.write("\n" +  json_object)
     
     print("------------------------------------------------------")

### 2023-04-27 data_cartography.py
stuff

@@ -1,0 +1,81 @@
+# in this file I am gonna recreate the data cartography repository.
+
+import datasets
+from transformers import AutoTokenizer, AutoModelForSequenceClassification, \
+    AutoModelForQuestionAnswering, Trainer, TrainingArguments, HfArgumentParser
+from helpers import prepare_dataset_nli, prepare_train_dataset_qa, \
+    prepare_validation_dataset_qa, QuestionAnsweringTrainer, compute_accuracy
+import os
+import json
+
+NUM_PREPROCESSING_WORKERS = 2
+
+# Loading the Data
+default_datasets = {'qa': ('squad',), 'nli': ('snli',)}
+dataset_id = ('snli',)
+eval_split = 'validation_matched' if dataset_id == ('glue', 'mnli') else 'validation'
+dataset = datasets.load_dataset(*dataset_id)
+model_path = './trained_model/'
+
+# TODO uncomment this
+# dataset = dataset.filter(lambda ex: ex['label'] != -1)
+
+train_dataset = dataset["train"]
+eval_dataset = dataset["validation"]
+test_dataset = dataset["test"]
+
+# example_indicies = range(10)
+# reduced_eval_dataset = train_dataset[example_indicies]
+
+tokenizer = AutoTokenizer.from_pretrained(model_path, use_fast=True)
+prepare_train_dataset = prepare_eval_dataset = \
+    lambda exs: prepare_dataset_nli(exs, tokenizer, 128)
+
+
+
+# Loading the model
+
+
+train_dataset_featurized = None
+eval_dataset_featurized = None
+
+eval_dataset_featurized = eval_dataset.map(
+    prepare_eval_dataset,
+    batched=True,
+    num_proc=NUM_PREPROCESSING_WORKERS,
+    remove_columns=eval_dataset.column_names
+)
+
+# eval_dataset_featureized = eval_dataset_featurized[0:10]
+
+
+model_classes = {'qa': AutoModelForQuestionAnswering,
+                    'nli': AutoModelForSequenceClassification}
+model_class = model_classes['nli']
+model = model_class.from_pretrained(model_path)
+
+
+
+# This function wraps the compute_metrics function, storing the model's predictions
+# so that they can be dumped along with the computed metrics
+compute_metrics = compute_accuracy
+# This function wraps the compute_metrics function, storing the model's predictions
+# so that they can be dumped along with the computed metrics
+eval_predictions = None
[…]

### 2023-04-27 paper/main.tex
progress

@@
 \usepackage{cite}
+\usepackage{graphicx}
 
@@
 
-\todo{You should describe modifications you have made to improve performances.
-We will evaluate based on how reasonable is your fix  -- if your fix and
-motivation does not link, your modificiation is unreasonable. For example,
-perhaps you tried to modify neural network training in a way that is totally
-unconnected to your stated goal, or your modification was erroneous. Describe
-the baseline model to be compared to your approach as well here. Again, clearly
-define all notations. Someone who is reading your report (with reasonable
-background, e.g., your classmates in this course) should have a reasonable idea
-how to implement it themselves. Describe hyperparameters, including how they
-are selected, if you have hyperparameters involved. }
+It is clear that the issue with the model is that it fails to understand
+ambiguous training examples. This is especially seen given that the model
+largely confuses, as seen earlier, neutral and entailment and instances of
+contradiction.
+
+\begin{figure}
+  \includegraphics[width=\linewidth]{./ambiguous.jpg}
+  \caption{Data Cartogram of 2000 training examples}
+  \label{}
+\end{figure}
+
+\subsection{Dataset Cartography}
+
+Training on labels that are ambiguous, that have high variablity, should in
+theory help with this issue, even more so than hard to learn labels. An
+ambiguous label is defined as a label that has high variance when predicted by
+different iterations of the model. For example, a model that is trained on 500
+iterations would predict the input differently than a model trained with 1000
+iterations or a model trained with 1500 iterations. Furthermore, the average
+confidence of that training example is not high or low.
+\cite{swayamdipta2020dataset} We reproduced the code to analyze data, and have
+plotted 500 training examples in figure 1.
+
+In order to deal with this, we first created a program that finds ambiguous
+examples. We found the variance of the training example based on its confidence
+of the gold label from a modeled trained on 500 iterations, 1000 iterations,
+1500 iterations, 2000 iterations, 2500 iterations, and 3000 iterations. If th
+variance was greater than 0.2, and the confidence of the model was in between
+0.6 and 0.7, the example was deemed ambiguous. According to \textit{Swayamdipta
+et al.}, training solely ambiguous examples leads to to greater performance
+than training on the whole dataset, only easy to learn examples, or only hard to
+learn examples. Thus it is logical to train on these examples.
+
+We pulled 1655 ambiguous samples from the training data and trained on those examples.
 

### 2023-04-27 cartography_analysis.py

@@ -1,0 +1,49 @@
+import json
+import numpy as np
+from scipy.special import softmax
+import matplotlib.pyplot as plt
+import seaborn as sns
+import pandas as pd
+
+results500 = open("./model500/eval_predictions.jsonl", "r")
+results1000 = open("./model1000/eval_predictions.jsonl", "r")
+results1500 = open("./model1500/eval_predictions.jsonl", "r")
+results2000 = open("./model2000/eval_predictions.jsonl", "r")
+results2500 = open("./model2500/eval_predictions.jsonl", "r")
+results3000 = open("./model3000/eval_predictions.jsonl", "r")
+results= open("./model/eval_predictions.jsonl", "r")
+
+results = [results500, results1000, results1500, results2000, results2500, results3000, results]
+num_results = len(results)
+num_lines = 50000
+
+confidence_array = np.zeros(shape=(num_results, num_lines))
+for i in range(num_results):
+    res = results[i]
+    for (j, line) in enumerate(res):
+        if j >= num_lines:
+            break
+        example = json.loads(line)
+        scores = np.array(example["predicted_scores"])
+        label = example["label"]
+        s_scores = softmax(scores)
+        gold_prob = s_scores[label]
+        confidence_array[i, j] = gold_prob
+
+# print(confidence_array)
+variance = np.std(confidence_array, axis=0)
+mean = np.mean(confidence_array, axis=0)
+
+df = pd.DataFrame({"confidence": mean, "variance": variance})
+df["ambigious"] = (df["variance"] > 0.15) & (df["confidence"] > 0.4) & (df["confidence"] < 0.8)
+# df["ambigious"] = np.all([variance > 0.2, mean > 0.6, mean < 0.7])
+# print(df.head())
+
+# sns.scatterplot(df, x="variance", y="confidence", hue=df["ambigious"])
+# plt.savefig("paper/ambiguous.jpg")
+# plt.show()
+
+l = df.index[df["ambigious"]].tolist()
+f = open("ambiguous_labels.txt", "a")
+for num in l:
+    f.write(str(num) + "\n")

### 2023-04-27 create_ambig_dataset.py

@@ -1,0 +1,16 @@
+import json
+
+f = open("ambiguous_labels.txt")
+m = open("model500/eval_predictions.jsonl")
+t = open("amiguous_examples.jsonl", "a")
+labels = []
+
+for label in f:
+    labels.append(int(label))
+
+for (i, line) in enumerate(m):
+    if(i in labels):
+        example = json.loads(line)
+        train_ex = {"premise" : example["premise"], "hypothesis" : example["hypothesis"], "label" : example["label"]}
+        json_object = json.dumps(train_ex) 
+        t.write(json_object + "\n")

### 2023-04-27 data_cartography.py

@@ -1,81 +1,0 @@
-# in this file I am gonna recreate the data cartography repository.
-
-import datasets
-from transformers import AutoTokenizer, AutoModelForSequenceClassification, \
-    AutoModelForQuestionAnswering, Trainer, TrainingArguments, HfArgumentParser
-from helpers import prepare_dataset_nli, prepare_train_dataset_qa, \
-    prepare_validation_dataset_qa, QuestionAnsweringTrainer, compute_accuracy
-import os
-import json
-
-NUM_PREPROCESSING_WORKERS = 2
-
-# Loading the Data
-default_datasets = {'qa': ('squad',), 'nli': ('snli',)}
-dataset_id = ('snli',)
-eval_split = 'validation_matched' if dataset_id == ('glue', 'mnli') else 'validation'
-dataset = datasets.load_dataset(*dataset_id)
-model_path = './trained_model/'
-
-# TODO uncomment this
-# dataset = dataset.filter(lambda ex: ex['label'] != -1)
-
-train_dataset = dataset["train"]
-eval_dataset = dataset["validation"]
-test_dataset = dataset["test"]
-
-# example_indicies = range(10)
-# reduced_eval_dataset = train_dataset[example_indicies]
-
-tokenizer = AutoTokenizer.from_pretrained(model_path, use_fast=True)
-prepare_train_dataset = prepare_eval_dataset = \
-    lambda exs: prepare_dataset_nli(exs, tokenizer, 128)
-
-
-
-# Loading the model
-
-
-train_dataset_featurized = None
-eval_dataset_featurized = None
-
-eval_dataset_featurized = eval_dataset.map(
-    prepare_eval_dataset,
-    batched=True,
-    num_proc=NUM_PREPROCESSING_WORKERS,
-    remove_columns=eval_dataset.column_names
-)
-
-# eval_dataset_featureized = eval_dataset_featurized[0:10]
-
-
-model_classes = {'qa': AutoModelForQuestionAnswering,
-                    'nli': AutoModelForSequenceClassification}
-model_class = model_classes['nli']
-model = model_class.from_pretrained(model_path)
-
-
-
-# This function wraps the compute_metrics function, storing the model's predictions
-# so that they can be dumped along with the computed metrics
-compute_metrics = compute_accuracy
-# This function wraps the compute_metrics function, storing the model's predictions
-# so that they can be dumped along with the computed metrics
-eval_predictions = None
[…]

### 2023-04-27 run_cartography_models.py

@@ -1,0 +1,204 @@
+import datasets
+from transformers import AutoTokenizer, AutoModelForSequenceClassification, \
+    AutoModelForQuestionAnswering, Trainer, TrainingArguments, HfArgumentParser
+from helpers import prepare_dataset_nli, prepare_train_dataset_qa, \
+    prepare_validation_dataset_qa, QuestionAnsweringTrainer, compute_accuracy
+import os
+import json
+
+NUM_PREPROCESSING_WORKERS = 2
+
+
+def main():
+    argp = HfArgumentParser(TrainingArguments)
+    # The HfArgumentParser object collects command-line arguments into an object (and provides default values for unspecified arguments).
+    # In particular, TrainingArguments has several keys that you'll need/want to specify (when you call run.py from the command line):
+    # --do_train
+    #     When included, this argument tells the script to train a model.
+    #     See docstrings for "--task" and "--dataset" for how the training dataset is selected.
+    # --do_eval
+    #     When included, this argument tells the script to evaluate the trained/loaded model on the validation split of the selected dataset.
+    # --per_device_train_batch_size <int, default=8>
+    #     This is the training batch size.
+    #     If you're running on GPU, you should try to make this as large as you can without getting CUDA out-of-memory errors.
+    #     For reference, with --max_length=128 and the default ELECTRA-small model, a batch size of 32 should fit in 4gb of GPU memory.
+    # --num_train_epochs <float, default=3.0>
+    #     How many passes to do through the training data.
+    # --output_dir <path>
+    #     Where to put the trained model checkpoint(s) and any eval predictions.
+    #     *This argument is required*.
+
+    argp.add_argument('--model', type=str,
+                      default='google/electra-small-discriminator',
+                      help="""This argument specifies the base model to fine-tune.
+        This should either be a HuggingFace model ID (see https://huggingface.co/models)
+        or a path to a saved model checkpoint (a folder containing config.json and pytorch_model.bin).""")
+    argp.add_argument('--task', type=str, choices=['nli', 'qa'], required=True,
+                      help="""This argument specifies which task to train/evaluate on.
+        Pass "nli" for natural language inference or "qa" for question answering.
+        By default, "nli" will use the SNLI dataset, and "qa" will use the SQuAD dataset.""")
+    argp.add_argument('--dataset', type=str, default=None,
[…]

### 2023-04-27 paper/main.tex
rough draft of paper done

@@
 \usepackage{graphicx}
+\usepackage[autostyle]{csquotes}
 
@@
 
-\date{}
+\date{April 27, 2023}
 
@@
 \begin{abstract}
-    \todo{write an abstract describing the core motivation and findings from your project}
+
+This paper describes the training and testing of an ELECTRA model for natural
+language inference using the Stanford Natural Language Inference Dataset. The
+model's performance is benchmarked using contrast sets, and data cartography is
+used to improve the model's accuracy. The results show that the base ELECTRA
+model performs well on the dataset, and extracting ambiguous examples from data
+cartography and training the model on them is an effective method for improving
+its performance for some aspects.
+
 \end{abstract}
@@
 
-\todo{This section should include three paragraphs. The first paragraph is to
-briefly describe task and data. The second paragraph should describe the
-results of your analysis, and the third paragraph describing your fix and main
-experimental take aways. }
+Natural language inference is the task of determining the truth value of a
+hypothesis given a premise. A hypothesis can either be true if it entails from
+the premise, false if it contradicts the premise, or neutral if a determination
+cannot be made either way. In this paper, train a model based on the Stanford
+Natural Language Inference Dataset (SNLI).
 
-The model that was used was \cite{clark2020electra}
+We find that the SNLI model has difficulty differenting between neutral labels,
+and labels of other classes. We also find that the model often is unable to
+contextualize and relate the semantics of different words that have the same
+meaning. We created a contrast set in order to test the model, and found that
+the performance of the model is significantly degraded in comparison to the
+validation set.
+
+In order to reduce ambiguity between words, we decided to train the model on
+ambiguous data - data with high variability and medium confidence. We found that
+training that although training the model on ambiguous data reduced the models
+overall performance, the performance of the model increased on adversarial data,
+which we dubbed contrast sets.
 
 \section{Task/Dataset/Model Description (15pt)}
+
+The model that was used was the ELECTRA model \cite{clark2020electra}. We are
+training  on the stanford NLI dataset \cite{bowman2015large}. The dataset has
+570000 training examples and 10000 validation examples. The baseline ELECTRA
+modeled was fine tuned to the SNLI training set and was trained with three
+epochs. We used the default loss function from the hugging face transformers
+library.
+
+For the fix, we took ambiguous data from the first 50000 training examples and
+trained on those for 15 epochs.
 
@@
   \includegraphics[width=\linewidth]{./ambiguous.jpg}
-  \caption{Data Cartogram of 2000 training examples}
+  \caption{Data Cartogram of 2000 training examples, orange ambiguous}
+  \label{}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=\linewidth]{./difficult.jpg}
+  \caption{Data Cartogram of 2000 training examples, orange difficult}
   \label{}
@@
 1500 iterations, 2000 iterations, 2500 iterations, and 3000 iterations. If th
-variance was greater than 0.2, and the confidence of the model was in between
-0.6 and 0.7, the example was deemed ambiguous. According to \textit{Swayamdipta
+variance was greater than 0.15, and the confidence of the model was in between
+0.4 and 0.8, the example was deemed ambiguous. According to \textit{Swayamdipta
 et al.}, training solely ambiguous examples leads to to greater performance
@@
 
-We pulled 1655 ambiguous samples from the training data and trained on those examples.
+We pulled 1655 ambiguous samples from the training data and trained on those
+examples. We trained for 15 epochs on ambiguous examples. As a test, we also
+trained solely on difficult to learn examples for five epochs, of which there
+were 6915, and benchmarked their performance.
 
@@
 
-\todo{
-Your writeup should address how effective is your fix, how broadly applicable
-is your fix, etc. Providing a single number (overall accuracy) is necessary but
-not sufficient here. For example, if your change made the model better on
-challenging NLI examples, you could try to quantify that on one or more slices
-of the data, give examples of predictions that are fixed, or even use model
-interpretation techniques to try to support claims about how your improved
-model is doing its ``reasoning.'' (You can look at the papers listed above to
-get a sense of how to do such fine-grained evaluation).  You should report
-results from a baseline approach (your initial trained model) as well as your
-``best'' method. If doing your own project, baselines such as majority class,
-random, or a linear classifier are important to see. \textbf{Ablations}: If you
-tried several things, analyze the contribution from each one. These should be
-\emph{minimal} changes to the same system; try running things with just one
-aspect different in order to assess how important that aspect is. This part of
-the report should be at least one page.}
+We evaluated the new model that was trained on hard to learn examples and
+independently trained on ambiguous examples. We found that models trained on
+ambiguous examples tended to do better with more difficult data, but tended to
+reduce the models overall average. For example, the accuracy of the model
+trained on ambiguous data and tested on the contrast set was 80\%, whereas the
+model that was not trained on ambiguous data had an accuracy of 77\%.
 
-% \todo{Describe the data you use, including how many examples are in the training, development, and test sets.
-Please also provide shallow statistics of your data, with at least the vocabulary size and instruction length.
-It is best to report all the statistics, including counts, in a table.
-Please describe how you compute statistics that can be computed in different ways (i.e., vocabulary size).
-Describe how you pre-process the data, including how you treat casing, tokenization, unknown words, and anything else that you did to the raw data before providing it to your model.
-If you are using a subset of the data, explain why and discuss tradeoffs (you will also need to back them with experiments).}
+Based on manual evaluation of the contrast set, there is some evidence that the
+model is able to reason better about differentiaing between entailment and
+contradictions (given that they have no errors) and in some instances better
+predicts neutrals. The following is a good example of this phenomenon:
 
-% \section{Implementation Details (3pt)}
+\begin{quote}
+premise: Under a blue sky with white clouds, a child reaches up to touch the propeller of a plane standing parked on a field of grass. \\
+hypothesis: A child is not reaching to touch the propeller out of curiousity.  \\
+label: neutral \\
+predicted: neutral
+\end{quote}
 
-% \todo{Briefly describe the implementation details.
-No need to copy details already specified in the assignment.
-Include any hyper-parameters the model has.
-If there are any optimizations that you introduced, this is the place to describe them.
-Did you have to do any special optimizations to make the method work in reasonable time/memory?
-Describe it here.}
+Unlike the the baseline model, which does not see the relationship that
+\textit{not} has in the sentence (negating curiousity, not touching), the model
+that is trained on ambiguous data is able to distinguish examples like these.
 
-% \section{}
+% Please add the following required packages to your document preamble:
+% \usepackagemultirow
+\begin{table}[]
+    \caption{Ambiguous Trained Model on Contrast Set}
+    \begin{tabular}{|ll|rrr|}
+        \hline
+        \multicolumn{2}{|l|}{\multirow{2}{*}{}}                           & \multicolumn{3}{l|}{\textbf{Predicted}}                                                              \\ \cline{3-5} 
+        \multicolumn{2}{|l|}{}                                            & \multicolumn{1}{l|}{\textit{0}} & \multicolumn{1}{l|}{\textit{1}}  & \multicolumn{1}{l|}{\textit{2}} \\ \hline
+        \multicolumn{1}{|l|}{\multirow{3}{*}{\textbf{True}}} & \textit{0} & \multicolumn{1}{r|}{90}         & \multicolumn{1}{r|}{7.692307692} & 0                               \\ \cline{2-5} 
+        \multicolumn{1}{|l|}{}                               & \textit{1} & \multicolumn{1}{r|}{10}         & \multicolumn{1}{r|}{61.53846154} & 30.76923077                     \\ \cline{2-5} 
+        \multicolumn{1}{|l|}{}                               & \textit{2} & \multicolumn{1}{r|}{0}          & \multicolumn{1}{r|}{7.692307692} & 92.30769231                     \\ \hline
+    \end{tabular}
+    \caption*{\textit{ 0=Entailment, 1=Neutral, 2=Contradiction }}
+\end{table}
+
+% Please add the following required packages to your document preamble:
+% \usepackage{multirow}
+\begin{table}[]
+    \caption{Ambiguous Trained Model on Validation Set}
+    \begin{tabular}{|ll|rrr|}
+        \hline
+        \multicolumn{2}{|l|}{\multirow{2}{*}{}}                           & \multicolumn{3}{l|}{\textbf{Predicted}}                                                             \\ \cline{3-5} 
+        \multicolumn{2}{|l|}{}                                            & \multicolumn{1}{l|}{\textit{0}} & \multicolumn{1}{l|}{\textit{1}} & \multicolumn{1}{l|}{\textit{2}} \\ \hline
+        \multicolumn{1}{|l|}{\multirow{3}{*}{\textbf{True}}} & \textit{0} & \multicolumn{1}{r|}{83.53}      & \multicolumn{1}{r|}{9.79}       & 7.04                            \\ \cline{2-5} 
+        \multicolumn{1}{|l|}{}                               & \textit{1} & \multicolumn{1}{r|}{6.39}       & \multicolumn{1}{r|}{79.75}      & 13.48                           \\ \cline{2-5} 
+        \multicolumn{1}{|l|}{}                               & \textit{2} & \multicolumn{1}{r|}{2.70}       & \multicolumn{1}{r|}{7.48}       & 89.871                          \\ \hline
+    \end{tabular}
+    \caption*{\textit{ 0=Entailment, 1=Neutral, 2=Contradiction }}
+\end{table}
+
+However, the model that was trained on ambiguous data seemed to suffer on the
+overall benchmark of the evaluation set. The accuracy dropped from 88\% to 84\%
+on the validation set.
+
+The most surprising results were the difficult to learn examples. The accuracy
+of that model was significantly lower than base model, clocking in an accuracy
+of 13\%. Even a model with random guessing could have done better.
+
+Due to time constraints, more ambiguous labels could not be extracted, but it's
+certainly possible that if all 500000 training examples are indexed and
+ambiguous ones are found, that the model could have better performance than it
+currently does. Furthremore, it is clear that training only on diffult to learn
+examples degrades model performance significantly.
 
  \section{Related Work (5pt)}
-\todo{Briefly discuss prior research papers related to your approach. This will
-likely some papers in the project description document. How is your approach
-different from existing studies?}
-% \paragraph{Ablations (15pt)}
-
-% \todo{Please report your ablation studies on the development set.
-Make sure to ablate each feature set independently to show they contribute to your final result.
-If a feature set does not improve performance, please do not include it -- it doesn't help.
-You can still discuss it, but there is really no point in including it in your final model.
-Reporting the results must be done in a table, and each result must be discussed in the text.}
-
-% \section{Error Analysis (7pt)}
-
-% \todo{Qualitative analysis of selected failure examples.
-Show and discuss error examples from your development set.
-You must identify certain classes of errors and use the examples to illustrate them.
-This is often best to show in a table.}
+Zhang et al. \cite{zhang2022cartography} propose another algorithm that uses
+data cartography called \enquote{cartography active learning}. However, their
+approach differs slightly in that they actively select ambiguous and difficult
+examples \textit{during} training and then adjust the input of the training in
+real time to train on the ambiguous data. In contrast, our approach finds
+ambiguous and difficult data after an initial training period and then trains on
+that data.
 
 \section{Conclusion (5pt)}
-\todo{Brief conclusion summarizing findings from both numerical results and
-qualitative analysis.}
 
-\section*{(Optional) AI Assistance}
-
-\todo{If you have used any AI toolkit (either for writing assistance or code
-assistance) for your final project, please describe it here. }
+In this paper, we evaluate an ELECTRA based ML model on the Stanford Natural
+Language Inference dataset. We find that the model has difficulty
+differentiating between neutral labels and other classes and that it finds at
+times spurious correlations between the premise and hypothesis. We also find
+that the model suffers significant performance degradation when tested against a
+contrast set. To remedy this, we recommend finding ambiguous labels (labels with
+high variablity and medium conference) using data cartography, and train the
+model on those data points in addition to the training set. Although not a
+perfect method, training on ambiguous labels can help the model learn more
+difficult correlations.
 

### 2023-04-27 analyze_output.py

@@ -4,7 +4,7 @@ from prettytable import PrettyTable
 from tabulate import tabulate
 from sklearn.metrics import confusion_matrix, ConfusionMatrixDisplay
 
-f = open("./eval_output/eval_predictions.jsonl")
+f = open("./ambiguous_only_validation/eval_predictions.jsonl")
 
 
 # {"premise": "Two women are embracing while holding to go packages.", 
@@ -16,11 +16,11 @@ f = open("./eval_output/eval_predictions.jsonl")
 mispredict = 0
 i = -1
 
-print("reference")
-print("0:", "Entailment")
-print("1:", "Neutral")
-print("2:", "Contradiction")
-print()
+# print("reference")
+# print("0:", "Entailment")
+# print("1:", "Neutral")
+# print("2:", "Contradiction")
+# print()
 
 y_true = []
 y_predict = []
@@ -41,15 +41,20 @@ for (i, line) in enumerate(f):
         # print("predicted:", example["predicted_label"])
         # print()
 
-print(mispredict, "/", i, "mispreditions")
+# print(mispredict, "/", i, "mispredictions")
 
 # Confusion Matrix
 cm = confusion_matrix(y_true=y_true, y_pred=y_predict)
-print(cm)
+# print(cm)
 
 # Confusion Matrix that prints out percentages
 np.set_printoptions(precision=2)
 percent = cm / cm.astype(np.float32).sum(axis=1) * 100
 table = tabulate(percent)
-print(table)
+# print(table)
 
+# print("---------------------")
+for row in percent:
+    for col in row:
+        print(col, end='\t')
+    print()

### 2023-04-27 cartography_analysis.py

@@ -36,14 +36,14 @@ mean = np.mean(confidence_array, axis=0)
 
 df = pd.DataFrame({"confidence": mean, "variance": variance})
 df["ambigious"] = (df["variance"] > 0.15) & (df["confidence"] > 0.4) & (df["confidence"] < 0.8)
-# df["ambigious"] = np.all([variance > 0.2, mean > 0.6, mean < 0.7])
-# print(df.head())
+df["difficult"] = (df["confidence"] < 0.5)
 
-# sns.scatterplot(df, x="variance", y="confidence", hue=df["ambigious"])
-# plt.savefig("paper/ambiguous.jpg")
+# sns.color_palette("flare", as_cmap=True)
+# sns.scatterplot(df, x="variance", y="confidence", hue=df["difficult"])
+# plt.savefig("paper/difficult.jpg")
 # plt.show()
 
-l = df.index[df["ambigious"]].tolist()
-f = open("ambiguous_labels.txt", "a")
+l = df.index[df["difficult"]].tolist()
+f = open("difficult_labels.txt", "a")
 for num in l:
     f.write(str(num) + "\n")

### 2023-04-27 create_ambig_dataset.py

@@ -1,8 +1,8 @@
 import json
 
-f = open("ambiguous_labels.txt")
+f = open("difficult_labels.txt")
 m = open("model500/eval_predictions.jsonl")
-t = open("amiguous_examples.jsonl", "a")
+t = open("difficult_examples.jsonl", "a")
 labels = []
 
 for label in f:

### 2023-04-27 paper/main.tex
some changes, time to review the paper lmao

@@
 
-For the fix, we took ambiguous data from the first 50000 training examples and
-trained on those for 15 epochs.
-
-\todo{Describe the task/dataset/model you are working on. Clearly define your
-task with mathematical notations. Describe your learning algorithm. You must
-formally specify the loss function, stopping criteria, training data, etc used
-for the model you are analyzing in the next section. Remember, every notation
-you use must be defined.}
+For the fix, we took 1655 ambiguous data examples from the first 50000 training
+examples and trained on those for 15 epochs on top of the original model. We
+also trained 6915 difficult to learn examples for 5 epochs on top of the
+baseline model and evaluate that.
 

### 2023-04-27 paper/main.tex
final paper

@@
 used to improve the model's accuracy. The results show that the base ELECTRA
-model performs well on the dataset, and extracting ambiguous examples from data
-cartography and training the model on them is an effective method for improving
-its performance for some aspects.
+model performs well on the dataset, and extracting \enquote{ambiguous} examples
+from data cartography and training the model on them is an effective method for
+improving its performance for some aspects.  We also found that
+\enquote{difficult to learn} training examples vastly reduce model performance,
+and suggest removing not having them in the training data.
 
@@
 training that although training the model on ambiguous data reduced the models
-overall performance, the performance of the model increased on adversarial data,
-which we dubbed contrast sets.
+overall performance by about 3\%, the performance of the model increased on
+adversarial data, which in this case are contrast sets, by 3\% as well. We also
+found that \enquote{difficult to learn} examples significantly lowered accuracy,
+and belive model performance can be improved with them removed.
 
@@
 
-The model that was used was the ELECTRA model \cite{clark2020electra}. We are
-training  on the stanford NLI dataset \cite{bowman2015large}. The dataset has
-570000 training examples and 10000 validation examples. The baseline ELECTRA
-modeled was fine tuned to the SNLI training set and was trained with three
-epochs. We used the default loss function from the hugging face transformers
-library.
+The model that was used was the ELECTRA model \cite{clark2020electra}, a
+transformer based on BERT. We are training  on the stanford NLI dataset
+\cite{bowman2015large}. The dataset has 570000 training examples and 10000
+validation examples. The baseline ELECTRA model was fine tuned to the SNLI
+training set and was trained with three epochs. We used the default loss
+function from the hugging face transformers library.
 
@@
 also trained 6915 difficult to learn examples for 5 epochs on top of the
-baseline model and evaluate that.
+baseline model and evaluate that. Our evaluation was accuracy for each of our
+different test/validation sets as well as a confusion matrix that describes the
+distribution of the data.
 
@@
 \subsection{Confusion Matrix Analysis}
+
 In general, the model seems to do well with Separating Entailment and
-Contradiction. For example, it only incorrectly identifies 2.47\% of entailment
-examples as contradiction and 2.37\% of contradiction examples as entailement.
+Contradiction. As can be seen in Table 2, it only incorrectly identifies 2.47\%
+of entailment examples as contradiction and 2.37\% of contradiction examples as
+entailement.
 
@@
 entailment and neutral, as well as contradiction and neutral. For example, it
-mispredicted over 7\% of entailments and contradictions as neutrals. As long as 
-mispredicting over 6\% of neutrals as entailments and over 7\% of of neutrals as
-contradictions.
+mispredicted over 7\% of entailments and contradictions as neutrals.
+Furthermore, it also mispredicts over 6\% of neutrals as entailments and over
+7\% of of neutrals as contradictions.
 
@@
 and instead makes assumption that they are different, thus leading it to think
-they are contradictions.
+they are contradictions. Further examples of this phenomenon can be seen in
+Table 4, which gives examples.
 
@@
 contrast sets} \cite{gardner2020evaluating}, Gardner et al. explain the concept
-of a \textit{contrast set}. A contrast set is a small but meaninful alteration
+of a \textit{contrast set}. A contrast set is a small but meaningful alteration
 in the input data that typically leads to a different gold label.
@@
 In order to test if the model is finding artifacts in the data or genuinely
-evaluating it, we annotated 35 different examples in the validation data that
+evaluating it, we hand annotated 35 different examples in the validation data that
 the model predicted correctly, and then tested to see if the model was able to
@@
 The results of the model showed a significant dip in accuracy, with accuracy
-going down to 77\% from 89\%. The model mispredicted 8/35 examples. One pattern
-that emerged from the mispredictions is that the model fails to see correlations
-between various words and also fails to understand the relationships between
-various words. The following is an example of this phenomenon: \\
+going down to 77\% from initially predicting them all correctly. Not to mention,
+the statistic is significantly lower than the model baseline of 88\% The model
+mispredicted 8/35 examples. One pattern that emerged from the mispredictions is
+that the model fails to see correlations between various words and also fails to
+understand the relationships between various words. The following is an example
+of this phenomenon: \\
 
@@
 ambiguous training examples. This is especially seen given that the model
-largely confuses, as seen earlier, neutral and entailment and instances of
-contradiction.
+largely confuses the semantics of various words that otherwise have similar
+meanings.
 
@@
 
+We believe that dataset cartography is a good fix to the problem. This is
+because ambiguous training examples typically map to \enquote{hard} problems,
+where as \enquote{hard to learn examples} typically map to training data with
+errors. Therefore, we recommend training without \enquote{hard to learn}
+examples and weighting \enquote{ambiguous} examples greater.
+
 In order to deal with this, we first created a program that finds ambiguous
@@
 of the gold label from a modeled trained on 500 iterations, 1000 iterations,
-1500 iterations, 2000 iterations, 2500 iterations, and 3000 iterations. If th
+1500 iterations, 2000 iterations, 2500 iterations, and 3000 iterations. If the
 variance was greater than 0.15, and the confidence of the model was in between
-0.4 and 0.8, the example was deemed ambiguous. According to \textit{Swayamdipta
-et al.}, training solely ambiguous examples leads to to greater performance
-than training on the whole dataset, only easy to learn examples, or only hard to
+0.4 and 0.8, the example was deemed ambiguous. We further deemed labels with
+confidence below 0.5 as \enquote{hard to learn} According to \textit{Swayamdipta
+et al.}, training solely ambiguous examples leads to to greater performance than
+training on the whole dataset, only easy to learn examples, or only hard to
 learn examples. Thus it is logical to train on these examples.
@@
 \section{Evaluating Your Fix (25pt)}
+
+\subsection{Ambiguous Examples}
 
@@
 overall benchmark of the evaluation set. The accuracy dropped from 88\% to 84\%
-on the validation set.
+on the validation set. Table 5 and 6 show the accuracy of the ambiguously
+trained model on the contrast and validation set, and shows that there is
+improvement in contrast set but some regression in the validation set.
 
-The most surprising results were the difficult to learn examples. The accuracy
-of that model was significantly lower than base model, clocking in an accuracy
-of 13\%. Even a model with random guessing could have done better.
+\subsection{Difficult to learn examples}
+
+Unsurprisingly, difficult to learn examples had accuracies that were
+significantly lower than the base model. However, it was surpring the extent to
+which they degraded the performance of the model. Like the ambiguous examples,
+the difficult to learn examples were trained directly on top of the base model,
+so the model was already partially fine tuned. Nevertheless, the 6915 training
+examples over 5 epochs degraded the model performance a very large amount. The
+accuracy of that model was significantly lower than base model, clocking in an
+accuracy of 13\%. Even a model with random guessing could have done better.
+Because of this, we highly suggest removing difficult to learn examples from the
+training data.
 
@@
 perfect method, training on ambiguous labels can help the model learn more
-difficult correlations.
+difficult correlations. We also suggest removing \enquote{difficult to learn}
+training examples, as they hurt model performance and are likely annotator
+errors. Dataset cartography overall is a highly effective method eliminating
+unwanted training examples or weight wanted training examples.
 
